---
getwd(title: "ML project"
output: html_document
date: "2025-02-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/Users/cindyyu/Desktop/CS4220")
library(readxl)
library(tidyverse)
library(caTools)

#random forest
library(randomForest)
library(caret)
library(VIM)

library(mice)
```

#READING IN FILTER PREDICTIONS
```{r}
real1_df <- read.table("snv-parse-real1.txt", header = T)
real2.1_df <- read.table("snv-parse-real2_part1.txt", header = T)
syn1_df <- read.table("snv-parse-syn1.txt", header = T)
syn2_df <- read.table("snv-parse-syn2.txt", header = T)
syn3_df <- read.table("snv-parse-syn3.txt", header = T)
syn4_df <- read.table("snv-parse-syn4.txt", header = T)
syn5_df <- read.table("snv-parse-syn5.txt", header = T)

all_df <- rbind(real1_df, real2.1_df, syn1_df, syn2_df, syn3_df, syn4_df, syn5_df)
rm(real1_df, real2.1_df, syn1_df, syn2_df, syn3_df, syn4_df, syn5_df)
```

#READING IN TRUTHS

```{r}
#add headers: chromosome #, genomic start, genomic stop
real1_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/real1/real1_truth.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
real2.1_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/real2_part1/real2_truth_chr1to5.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
syn1_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/syn1/syn1_truth.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
syn2_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/syn2/syn2_truth.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
syn3_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/syn3/syn3_truth.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
syn4_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/syn4/syn4_truth.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
syn5_truth <- as.data.frame(read.table("/Users/cindyyu/Desktop/CS4220/Data/syn5/syn5_truth.bed",
                                header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
alltruth_df <- rbind(real1_truth, real2.1_truth, syn1_truth, syn2_truth, syn3_truth, syn4_truth, syn5_truth)

names(alltruth_df) <- c("Chr", "START_POS_REF", "END_POS_REF")

write.table(alltruth_df, "all_truths.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)
rm(real1_truth, real2.1_truth, syn1_truth, syn2_truth, syn3_truth, syn4_truth, syn5_truth)
```


```{r}
alltruth_df <- alltruth_df %>% mutate(TRUTH = TRUE)

df3 <- all_df %>% 
  left_join(alltruth_df, by=c('Chr','START_POS_REF', 'END_POS_REF'))

df3$TRUTH <- replace_na(df3$TRUTH, FALSE)

```



#DATA IMPUTATION
```{r}
#replace NAs w MEAN
df3_narm <- df3 %>% mutate_all(~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x)) 

#grouped replacing NAs w mean 
df_groupmean <- df3 %>%
  group_by(FILTER_Mutect2) %>%
  mutate(m2_MQ = ifelse(is.na(m2_MQ), mean(m2_MQ, na.rm = TRUE), m2_MQ)) |>
  ungroup() %>%
  group_by(FILTER_Freebayes) %>%
  mutate(f_MQMR = ifelse(is.na(f_MQMR), mean(f_MQMR, na.rm = TRUE), f_MQMR)) %>%
  ungroup() %>%
  group_by(FILTER_Vardict) %>%
  mutate(vd_SSF = ifelse(is.na(vd_SSF), mean(vd_SSF, na.rm = TRUE), vd_SSF)) %>%
  ungroup() %>%
  group_by(FILTER_Vardict) %>%
  mutate(vd_MSI = ifelse(is.na(vd_MSI), mean(vd_MSI, na.rm = TRUE), vd_MSI)) %>%
  ungroup() %>%
  group_by(FILTER_Varscan) %>%
  mutate(vs_SSC = ifelse(is.na(vs_SSC), mean(vd_MSI, na.rm = TRUE), vs_SSC)) %>%
  ungroup() %>%
  group_by(FILTER_Varscan) %>%
  mutate(vs_SPV = ifelse(is.na(vs_SPV), mean(vs_SPV, na.rm = TRUE), vs_SPV)) %>%
  ungroup()

df_groupmedian <- df3 %>%
  group_by(FILTER_Mutect2) %>%
  mutate(m2_MQ = ifelse(is.na(m2_MQ), median(m2_MQ, na.rm = TRUE), m2_MQ)) |>
  ungroup() %>%
  group_by(FILTER_Freebayes) %>%
  mutate(f_MQMR = ifelse(is.na(f_MQMR), median(f_MQMR, na.rm = TRUE), f_MQMR)) %>%
  ungroup() %>%
  group_by(FILTER_Vardict) %>%
  mutate(vd_SSF = ifelse(is.na(vd_SSF), median(vd_SSF, na.rm = TRUE), vd_SSF)) %>%
  ungroup() %>%
  group_by(FILTER_Vardict) %>%
  mutate(vd_MSI = ifelse(is.na(vd_MSI), median(vd_MSI, na.rm = TRUE), vd_MSI)) %>%
  ungroup() %>%
  group_by(FILTER_Varscan) %>%
  mutate(vs_SSC = ifelse(is.na(vs_SSC), median(vd_MSI, na.rm = TRUE), vs_SSC)) %>%
  ungroup() %>%
  group_by(FILTER_Varscan) %>%
  mutate(vs_SPV = ifelse(is.na(vs_SPV), median(vs_SPV, na.rm = TRUE), vs_SPV)) %>%
  ungroup()

```

#MICE

```{r}
md.pattern(df3)

library(VIM)
aggr_plot <- aggr(df3, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(df3), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))

```

```{r}
#missForest
install.packages("missForest")
library(missForest)

df3_features <- df3[9:19]
df3_features$FILTER_Mutect2 = as.factor(df3_features$FILTER_Mutect2)
df3_features$FILTER_Freebayes = as.factor(df3_features$FILTER_Freebayes)
df3_features$FILTER_Vardict = as.factor(df3_features$FILTER_Vardict)
df3_features$FILTER_Varscan = as.factor(df3_features$FILTER_Varscan)
df3_features$TRUTH = as.factor(df3_features$TRUTH)

#seed 10% missing values
df3.mis <- prodNA(df3_features, noNA = 0.1)
summary(df3.mis)

#impute missing values, using all parameters as default values
df3.imp <- missForest(df3.mis)

#check imputed values
df3.imp$ximp

#check imputation error
df3.imp$OOBerror

```


#MACHINE LEARNING 
```{r}
#remove chr and genomic start/stop

#splitting
#SOMEHOW REMOVE COLUMNS OF GENOME NUMBER AND CHR
split_vals <- sample.split(df3_narm $TRUTH, SplitRatio = 0.80)
train_set <- subset(df3_narm, split_vals == TRUE)
test_set <- subset(df3_narm, split_vals == FALSE)

train_set_split <- train_set[,-(1:8)]

#model
rpart(as.factor(TRUTH)~ ., data = train_set_split) -> mod_class
predict(mod_class, test_set, type = 'class') -> result_class
table(test_set$TRUTH, result_class)

#refitting predictions to be bed file
#REALIZED I SHOULDNT HAVE UPENDED TO DF3, BUT RATHER TEST SET!
predictions <- data.frame(Chr = test_set$Chr,
                          START_POS_REF = test_set$START_POS_REF,
                          END_POS_REF = test_set$END_POS_REF, 
                          TRUTH=result_class)

predictions <- predictions |>
  filter(TRUTH == TRUE) 

my_predictions <- predictions[, -4]

write.table(my_predictions, "my-predictions.bed",row.names = F,col.names = F, sep="\t", quote=FALSE) 


#change columsn to remove chromo number
#treat diff samples differently

```

#RANDOM FOREST ATTEMPT - CINDY
```{r}





----------------------------------------------------------------------------------------------------------------------

neha 

-------------------------------------------------------------------------------------------------------------------------


ALSOOOOO 

for knn: 

# Step 1: Install and load the VIM package if you haven't already
install.packages("VIM")
library(VIM)

# Step 2: Perform KNN imputation only on specific variables
data <- kNN(new_data, variable = c("EducationLevel", "DietPerception"), k = 5)

# Step 3: Verify that missing values have been imputed
summary(data)

---
getwd(title: "ML project"
output: html_document
date: "2025-02-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("C:/Users/nehar/OneDrive - Victoria University of Wellington - STUDENT/EXCHANGE/ACADEMICS/CS/FOR THE PROJECT")

install.packages("readxl", "tidyverse", "caTools")

install.packages("knitr") 


install.packages("dplyr")


library(dplyr)


library(readxl)
library(tidyverse)
library(caTools)
library(knitr)
```

#READING IN FILTER PREDICTIONS
```{r}
real1_df <- read.table("snv-parse-real1.txt", header = T)
#real2.1_df <- read.table("snv-parse-real2_part1.txt", header = T)
real2.1_df <- read.table("snv-parse-real2_part1.txt ", header = T)
real2.2_df <- read.table("snv-parse-real2_part2.txt", header = T)

syn1_df <- read.table("snv-parse-syn1.txt", header = T)
syn2_df <- read.table("snv-parse-syn2.txt", header = T)
syn3_df <- read.table("snv-parse-syn3.txt", header = T)
syn4_df <- read.table("snv-parse-syn4.txt", header = T)
syn5_df <- read.table("snv-parse-syn5.txt", header = T)

all_df <- rbind(real1_df, real2.2_df, real2.2_df, syn1_df, syn2_df, syn3_df, syn4_df, syn5_df)

head(all_df)

```


```{r}

#looking for NA's
summary(all_df)



```
#READING IN TRUTHS

```{r}

file.exists("C:/Users/nehar/OneDrive - Victoria University of Wellington - STUDENT/EXCHANGE/ACADEMICS/CS/real1_truth.bed")


```

```{r}


real1_truth <- as.data.frame(read.table("C:/Users/nehar/OneDrive - Victoria University of Wellington - STUDENT/EXCHANGE/ACADEMICS/CS/real1_truth.bed",
                                       header = FALSE, sep="\t", stringsAsFactors=FALSE, quote=""))



real2.1_truth <- as.data.frame(read.table("C:/Users/nehar/OneDrive - Victoria University of Wellington - STUDENT/EXCHANGE/ACADEMICS/CS/real2_truth_chr1to5.bed",
                                          header = FALSE, sep="\t", stringsAsFactors=FALSE, quote=""))



syn1_truth <- as.data.frame(read.table("C:/Users/nehar/OneDrive - Victoria University of Wellington - STUDENT/EXCHANGE/ACADEMICS/CS/syn1_truth.bed",
                                       header = FALSE, sep="\t", stringsAsFactors=FALSE, quote=""))


syn2_truth <- as.data.frame(read.table("C:/Users/nehar/OneDrive - Victoria University of Wellington - STUDENT/EXCHANGE/ACADEMICS/CS/syn2_truth.bed",
                                       header = FALSE, sep="\t", stringsAsFactors=FALSE, quote=""))


syn3_truth <- as.data.frame(read.table("C:/Users/nehar/OneDrive - Victoria University of Wellington - STUDENT/EXCHANGE/ACADEMICS/CS/syn3_truth.bed",
                                       header = FALSE, sep="\t", stringsAsFactors=FALSE, quote=""))


syn4_truth <- as.data.frame(read.table("C:/Users/nehar/OneDrive - Victoria University of Wellington - STUDENT/EXCHANGE/ACADEMICS/CS/syn4_truth.bed",
                                       header = FALSE, sep="\t", stringsAsFactors=FALSE, quote=""))


syn5_truth <- as.data.frame(read.table("C:/Users/nehar/OneDrive - Victoria University of Wellington - STUDENT/EXCHANGE/ACADEMICS/CS/syn5_truth.bed",
                                       header = FALSE, sep="\t", stringsAsFactors=FALSE, quote=""))



alltruth_df <- rbind(real1_truth, real2.1_truth, syn1_truth, syn2_truth, syn3_truth, syn4_truth, syn5_truth)


names(alltruth_df) <- c("Chr", "START_POS_REF", "END_POS_REF")




head(alltruth_df)


```

```{r}

#thought: should we weight the real information more?
#GET RID OF CORREXT X AND Y
#MAKE COLUMN ALL TRUE IN ALL TRUTH
alltruth_df <- alltruth_df %>% mutate(TRUTH = TRUE)


df3 <- all_df %>% 
  left_join(alltruth_df, by=c('Chr','START_POS_REF', 'END_POS_REF'))


all_df <- all_df %>%
  left_join(alltruth_df %>% mutate(correct = TRUE), 
            by = c("Chr", "START_POS_REF", "END_POS_REF")) 
 

all_df <- all_df %>%
  mutate(correct = any(alltruth_df$Chr == Chr & alltruth_df$START_POS_REF == START_POS_REF &
                         alltruth_df$END_POS_REF == END_POS_REF))

```



#CREATING BENCHMARKS 

## consensus between at least 2

```{r}
consensus_df <- all_df %>%
  mutate(two_plus = rowSums(select(., FILTER_Mutect2, FILTER_Freebayes, FILTER_Varscan, FILTER_Vardict)) >= 2)
```

handelling missing data :)

```{r}



all_df <- all_df %>%
  mutate(across(everything(), 
                ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))


summary(all_df)




```




EDA:

ok idk how helpful this one is: 

```{r}

  
  install.packages("tidyverse")
  
  
  install.packages("ggcorrplot")
  
  

  library(tidyverse)
  library(ggcorrplot)  # for correlation heatmap
  
 olumns
  numeric_cols <- all_df %>% 
    select(m2_MQ, f_MQMR, vs_SSC, vs_SPV, vd_SSF, vd_MSI)
  

  numeric_cols <- numeric_cols %>%
    mutate(across(everything(), ~ifelse(is.na(.), median(., na.rm = TRUE), .)))
  
 
 
  cor_matrix <- cor(numeric_cols, use = "pairwise.complete.obs")
  
  
  ggcorrplot(cor_matrix, lab = TRUE, type = "lower", title = "Correlation Matrix")
  

```

eda: na values

```{r}


library(ggplot2)


na_count <- sapply(all_df, function(x) sum(is.na(x)))
na_percentage <- sapply(all_df, function(x) mean(is.na(x)))


na_df <- data.frame(Column = names(na_percentage), NA_Percentage = na_percentage)
ggplot(na_df, aes(x = reorder(Column, NA_Percentage), y = NA_Percentage)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "NA Distribution in Columns", x = "Columns", y = "Percentage of NAs") +
  theme_minimal()


missing_indicators <- all_df %>%
  mutate(across(everything(), ~ ifelse(is.na(.), 1, 0), .names = "missing_{.col}"))


all_df_with_missing_indicators <- cbind(all_df, missing_indicators)

head(all_df_with_missing_indicators)



```



#MACHINE LEARNING 

```{r}


install.packages("caTools")


#update r before doing this one it had some major issues for me 

install.packages("randomForest")

library(randomForest)
library(caTools)


all_df <- all_df %>%
  mutate(across(everything(), 
                ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))



set.seed(123)

#this is givign me a few errors - fixing 

split <- sample.split(all_df$correct, SplitRatio = 0.8)
training_set <- subset(all_df, split == TRUE)
test_set <- subset(all_df, split == FALSE)


rf_model <- randomForest(correct ~ ., data = training_set, importance = TRUE)

test_predictions <- predict(rf_model, newdata = test_set)


confusion_matrix <- table(test_set$correct, test_predictions)
print(confusion_matrix)


importance(rf_model)
varImpPlot(rf_model)





```







